---
title: "ANALYSIS of CLINICAL ACTIVITY"
author: "Dr. Andrea PECCATI"
date: "`r Sys.Date()`"
output:
  rmdformats::robobook:
    toc_depth: 3
    thumbnails: false
    lightbox: true
    gallery: false
    collapsed: false
---

```{=html}
<style>
@media print {
  /* 1. Hide menus, navigation bars, and logos */
  #sidebar, .book-summary, .book-header, .book-logo {
    display: none !important;
  }

  /* 2. Unlock the main container to allow page breaks */
  html, body, .book, .book-body, .page-wrapper, .page-inner, .section {
    display: block !important;
    height: auto !important;
    overflow: visible !important;
    position: static !important;
    margin: 0 !important;
    padding: 0 !important;
  }

  /* 3. Remove the rightward offset created by the sidebar */
  .book-body {
    left: 0 !important;
  }

  /* 4. Prevent charts/images from being split across two pages */
  img, canvas, .ggplot2 {
    page-break-inside: avoid !important;
  }
}
</style>
```

```{r setup, include = FALSE}
knitr::opts_chunk$set(
	fig.align = "center",
	fig.height = 6,
	fig.width = 8,
	message = FALSE,
	warning = FALSE,
	dpi = 100,
	out.height = "600px",
	out.width = "800px"
)
load("clinical_analysis.RData")
library(kableExtra)
library(tidyverse)

# Create plot theme
  my_plot_theme <- theme_minimal() +
    theme(
      plot.title = element_text(face = "bold", size = 22, hjust = 0.5, margin = margin(t = 10, b = 20)),
      axis.title = element_text(face = "bold", size = 12),
      axis.text  = element_text(face = "bold", size = 12),
      axis.line  = element_line(color = "black", size = 0.5),
      axis.ticks = element_line(color = "black")
    )
```

<a href="" onclick="window.print()" class="print-button" name="top">

Print report

</a>

```{=html}
<style>
.print-button {
  font-weight: bold;
  text-align: center;
}
@media print {
  .print-button {
    display: none !important;
  }
}
</style>
```

\
The following objects were **exported into .RData format** from the `clinical_analysis.R` file to be imported in this report:

-   the `check_missing` function

-   the `clinical_total` dataframe

-   the `clinical_stats_per_year_wide` dataframe

-   the `clinical_stats_per_type_long` dataframe

-   the `clinical_stats_per_time_wide` dataframe

## 1 - DATA CHECK

Before data analysis, a quick **data check was performed to search for NA values** (amount and %).

```{r}
check_missing(clinical_total) |>
  rename(
    "Variable" = variable,
    "NA values" = missing,
    "%" = pct,
  ) |>
  kbl(align="c")|>
  kable_styling(
    bootstrap_options = "striped", 
    full_width = FALSE,      
    position = "center"
  ) |>
  column_spec(1:3, width = "120px")
```

## 2 - PATIENTS per YEAR

This is the report of the total amount and running total of **patients visited between `r min(clinical_stats_per_year_wide$year)` and `r max(clinical_stats_per_year_wide$year)`, [grouped by year]{.underline}.**

### 2.1 - Table

```{r}
clinical_stats_per_year_wide |> 
  select(year, patients) |>
  mutate(cumulative = cumsum(patients)) |>
  rename(
    "Year" = year,
    "Patients" = patients,
    "Running" = cumulative
  ) |>
  kbl(align="c")|>
  kable_styling(
    bootstrap_options = "striped", 
    full_width = FALSE,      
    position = "center"
  ) |>
  column_spec(1:3, width = "120px")
```

### 2.2 - Plot

```{r}
ggplot(
  data = clinical_stats_per_year_wide,
  aes(x = year)
) +
geom_bar(
  aes(y = patients),
  stat = "identity",
  fill = "lightgrey",
  color = "black",
  alpha = 0.6
) +
geom_text(
  aes(y = patients, label = patients),
  vjust = -0.5,
  fontface = "bold",
  size = 4
) +
scale_x_continuous(
  name ="",
  breaks = seq(
    min(clinical_stats_per_year_wide$year),
    max(clinical_stats_per_year_wide$year),
    by = 1
  )
) +
scale_y_continuous(
  name = "",
  expand = expansion(mult = c(0, 0.05))
) +
labs(
  title = paste0(
    "PATIENTS per YEAR (",
    min(clinical_stats_per_year_wide$year),
    " - ",
    max(clinical_stats_per_year_wide$year),
    ")"
  )
) +
my_plot_theme
```

## 3 - VISITS per YEAR and SERVICE

This is the report of the total amount and running total of **visits performed between `r min(clinical_stats_per_year_wide$year)` and `r max(clinical_stats_per_year_wide$year)`, [grouped by year ans service:]{.underline}**

-   First visit (new patients)

-   Follow-up visits

-   Hip US for DDH screening

-   Miscellaneous (e.g. cast removal)

### 3.1 - Table

```{r}
totals_row <- clinical_stats_per_year_wide |>
  summarise(
    year = "",
    visort01 = sum(visort01, na.rm = TRUE),
    visort02 = sum(visort02, na.rm = TRUE),
    ectbac = sum(ectbac, na.rm = TRUE),
    misc = sum(misc, na.rm = TRUE),
    patients = sum(patients, na.rm = TRUE)
  )

clinical_stats_per_year_wide |>
  select(year, visort01, visort02, ectbac, misc, patients) |>
  mutate(year = as.character(year)) |>
  bind_rows(totals_row) |>
  mutate(
    running = ifelse(year == "", "", cumsum(patients))
  ) |>
  rename(
    "Year" = year,
    "First visit" = visort01,
    "Follow-up" = visort02,
    "Hip US" = ectbac,
    "Misc" = misc,
    "Patients" = patients,
    "Running" = running
  ) |>
  kbl(align="c")|>
  kable_styling(
    bootstrap_options = "striped", 
    full_width = FALSE,      
    position = "center"
  ) |>
  column_spec(1:6, width = "100px")|>
  row_spec(
    nrow(clinical_stats_per_year_wide) + 1,
    bold = TRUE, 
    extra_css = "border-top: 2px solid black;"
  )
```

### 3.2 - Plot

```{r}
# 1. Convert clinical_stats_per_year from wide to long format
clinical_stats_per_year_long <- clinical_stats_per_year_wide |>
  pivot_longer(
    cols = c("visort01", "visort02", "ectbac", "misc"),
    names_to = "category",
    values_to = "value"
  ) |>
  mutate(category = factor(
    category,
    levels = c("misc", "ectbac", "visort02", "visort01")
    )
  )

# 2. Plot
ggplot(
  data = clinical_stats_per_year_long,
  aes(
    x = factor(year),
    y = value,
    fill = category)
) +
geom_bar(
  stat = "identity",
  position = "stack",
  color = "black"
) +
geom_text(
  aes(label = ifelse(value > 50, as.character(value), "")),
  position = position_stack(vjust = 0.5),
  fontface = "bold",
  size = 4
) +
scale_fill_manual(
  values = c(
    "visort01" = "grey95",
    "visort02" = "grey75", 
    "ectbac"   = "grey55", 
    "misc"     = "grey35"
  ),
  breaks = c(
    "visort01",
    "visort02",
    "ectbac",
    "misc"
  ),
  labels = c(
    "First visit",
    "Follow-up",
    "Hip US",
    "Misc"
  )
) + 
scale_y_continuous(
  name = "",
  expand = expansion(mult = c(0, 0.05))
) +
labs(
  title = paste0(
    "VISITS per YEAR and SERVICE (",
    min(clinical_stats_per_year_long$year),
    " - ",
    max(clinical_stats_per_year_long$year),
    ")"
  ),
  caption = "Only values above 50 are shown in bars",
  x = NULL,
  y = NULL,
  fill = NULL
) +
theme_minimal() +
my_plot_theme +
theme(
  legend.position = "bottom",
  axis.title.x = element_blank(),
  axis.title.y = element_blank()
)
```

## 4 - VISITS per YEAR and TYPE

This is the report of the total amount and running total of **visits performed** **between `r min(clinical_stats_per_year_wide$year)` and `r max(clinical_stats_per_year_wide$year)`, [grouped by year and type:]{.underline}**

-   National Health System

-   Private practice (no insurance)

-   Insurance

### 4.1 - Table

```{r}
clinical_stats_per_type_wide <- clinical_stats_per_type_long |>
  pivot_wider(
    names_from = type_cat,
    values_from = patients,
    values_fill = 0
  ) |>
  ungroup(year)

totals_row <- clinical_stats_per_type_wide |>
  summarise(
    year = "",
    NHS = sum(as.numeric(NHS), na.rm = TRUE),
    Insurance = sum(as.numeric(Insurance), na.rm = TRUE),
    Private = sum(as.numeric(Private), na.rm = TRUE),
    patients = NHS + Insurance + Private
  )

clinical_stats_per_type_wide |>
  mutate(
    year = as.character(year),
    patients = NHS + Insurance + Private
    ) |>
  bind_rows(totals_row) |>
  mutate(
    running = ifelse(year == "", "", cumsum(patients))
  ) |>
  rename(
    "Year" = year,
    "NHS" = NHS,
    "Insurance" = Insurance,
    "Private" = Private,
    "Patients" = patients,
    "Running" = running
  ) |>
  kbl(align="c")|>
  kable_styling(
    bootstrap_options = "striped", 
    full_width = FALSE,      
    position = "center"
  ) |>
  column_spec(1:6, width = "100px") |>
  row_spec(
    nrow(clinical_stats_per_type_wide) + 1,
    bold = TRUE, 
    extra_css = "border-top: 2px solid black;"
  )
```

### 4.2 - Plot

```{r}
clinical_stats_per_type_long$type_cat <- factor(
  clinical_stats_per_type_long$type_cat, 
  levels = c(
      "NHS",
      "Insurance",
      "Private"
    )
)

ggplot(
  data = clinical_stats_per_type_long,
  aes(
    x = factor(year),
    y = patients,
    fill = type_cat)
) +
geom_bar(
  stat = "identity",
  position = "stack",
  color = "black"
) +
geom_text(
  aes(label = ifelse(patients > 50, as.character(patients), "")),
  position = position_stack(vjust = 0.5),
  fontface = "bold",
  size = 4
) +
scale_fill_manual(
  values = c(
    "NHS" = "grey95",
    "Insurance" = "grey75",
    "Private" = "grey55"
  ),
  breaks = c(
    "NHS",
    "Insurance",
    "Private"
  ), 
  labels = c(
    "National Health System",
    "Private practice (Insurance)",
    "Private practice (no Insurance)"
  )
) +
scale_y_continuous(
  name = "",
  expand = expansion(mult = c(0, 0.05))
) +
labs(
  title = paste0(
    "VISITS per YEAR and TYPE (",
    min(clinical_stats_per_type_long$year),
    " - ",
    max(clinical_stats_per_type_long$year),
    ")"
  ),
  caption = "Only values above 50 are shown in bars",
  x = NULL,
  y = NULL,
  fill = NULL
) +
theme_minimal() +
my_plot_theme +
theme(
  legend.position = "bottom",
  axis.title.x = element_blank(),
  axis.title.y = element_blank()
)
```

## 5 - PATIENTS per TIME SLOT

This is the report of the total amount and running total of **visits performed, [grouped by time slot]{.underline}** (split into 2-hours slots from 08:00 to 20:00).

### 5.1 - Table

```{r}
clinical_stats_per_time_wide |>
  mutate(
    pct = sprintf("%.2f%%", patients/sum(patients)*100),
    cumulate = cumsum(patients),
    cumulate_pct = sprintf("%.2f%%",cumsum(patients/sum(patients)*100))
  )|>
  rename(
    "Time range" = time_cat,
    "Patients" = patients,
    "%" =  pct,
    "Running" = cumulate,
    "Runinng %" = cumulate_pct
  ) |>
  kbl(align="c")|>
  kable_styling(
    bootstrap_options = "striped", 
    full_width = FALSE,      
    position = "center"
  ) |>
  column_spec(1, width = "120px") |>
  column_spec(2:5, width = "100px")
```

### 5.2 - Plot

```{r}
ggplot(
  data = clinical_stats_per_time_wide,
  aes(x = time_cat)
) +
geom_bar(
  aes(y = patients),
  stat = "identity",
  fill = "lightgrey",
  color = "black",
  alpha = 0.6
) +
geom_text(
  aes(y = patients, label = patients),
  vjust = -0.5,
  fontface = "bold",
  size = 5
) +
scale_y_continuous(
  name = "",
  expand = expansion(mult = c(0, 0.05))
) +
labs(
  title = "PATIENTS per TIME SLOT",
  x = NULL
) +
my_plot_theme
```
